Hydra Battles and Cie

### 2021-07-30 doc/part-hydras.tex
converting chapter 2 to alectryon (continued)

\index{hydras}{Library Hydra!Predicates!round\_n}
\begin{Coqsrc}
Definition round_n n h h' := R1 h h' \/ R2 n h h'.  
\end{Coqsrc}

@@

\index{hydras}{Library Hydra!Predicates!round}
\label{sect:infix-round}
\begin{Coqsrc}
Definition round h h' := exists n,  round_n n h h'.

Infix "-1->" := round (at level 60).
\end{Coqsrc}

@@

\begin{Coqsrc}
Definition round_plus := clos_trans_1n Hydra round.
Infix "-+->" := rounds (at level 60).

Definition round_star h h' := h = h' \/ round_plus h h'.
Infix "-*->" := round_star (at level 60).
\end{Coqsrc}

@@

\begin{exercise}
Prove the following lemma:

\begin{Coqsrc}
Lemma rounds_height : forall h h', 
   h -+-> h' -> height h' <= height h.  
\end{Coqsrc}

@@

\begin{Coqsrc}
Definition dep_round_t := nat -> Hydra -> Hydra -> Prop.

Class Battle :=  
    { battle_rel : dep_round_t;
      battle_inclusion : forall i h h', battle_rel i h h' -> round h h'}.
\end{Coqsrc}

@@

\begin{Coqsrc}
Program Instance free : Battle := Build_Battle (fun _  h h' => round h h') _.
\end{Coqsrc}

@@

\begin{Coqsrc}
Program Instance standard : Battle := (Build_Battle round_n _).
Next Obligation.
  now exists i.  
Defined.
\end{Coqsrc}

@@

\begin{Coqsrc}
Inductive battle (B:Battle) : nat -> Hydra -> nat -> Hydra -> Prop :=
| battle_1 : forall i h  h', 
      battle_rel   B i  h h' -> battle B i h (S i) h'
| battle_n : forall i h  j h' h'',  
      battle_rel  B i h h''  ->
      battle B (S i) h'' j h'  ->
      battle B i h j h'.
\end{Coqsrc}

The following property allows us to build battles out of smaller ones.

\begin{Coqsrc}
Lemma battle_trans {b:Battle} :
  forall i h j h', battle b i h j h' ->
        forall k h0,  battle b k h0 i h -> battle b k h0 j h'.
\end{Coqsrc}

@@

\begin{Coqsrc}
Definition hinit := hyd3 (hyd_mult head 3)  head head.  
\end{Coqsrc}

@@

\begin{Coqsrc}
Lemma L_0_2 : battle standard 0 hinit 2 (hyd1 h3).   
\end{Coqsrc}

@@

A first study with pencil and paper suggested us that, after three rounds, the hydra always looks like in figure~\vref{fig:hinit-plusn} (with a variable number of 
subtrees of height 1 or 0).
Thus, we introduce a few handy notations.

\begin{Coqsrc}
Notation h3 := (hyd_mult head 3).
Notation h2 := (hyd_mult head 2).
Notation h1 := (hyd1 head).

Definition hyd a b c := 
  node (hcons_mult h2  a
             (hcons_mult h1  b
                         (hcons_mult head c hnil))).
\end{Coqsrc}

@@

\begin{Coqsrc}
Lemma L_2_3 : battle standard 2 (hyd1 h3)  3 (hyd 3 0 0).

Lemma L_0_3 : battle standard 0 hinit 3 (hyd 3 0 0).
\end{Coqsrc}

@@

\begin{Coqsrc}
 Record state : Type :=
    mks {round: nat ; n2 : nat ; n1 : nat ; nh : nat}.
\end{Coqsrc}

@@

\begin{Coqsrc}
Definition next (s : state) :=
  match s with
  | mks round a b (S c) => mks (S round) a b c
  | mks round a (S b) 0 => mks (S round) a b (S round)
  | mks round (S a) 0 0 => mks (S round) a (S round) 0
  | _ => s
  end.
\end{Coqsrc}

@@

\begin{Coqsrc}
Fixpoint iterate {A:Type}(f : A -> A) (n: nat)(x:A) :=
  match n with
  | 0 => x
  | S p => f (iterate  f p x)
  end.
\end{Coqsrc}

@@

\begin{Coqsrc}
Definition test n := iterate next (n-3) (mks 3 3 0 0).

Compute test 3.
\end{Coqsrc}

\begin{Coqanswer}
      = {| round := 3; n2 := 3; n1 := 0; nh := 0 |}
     : state 
\end{Coqanswer}

\begin{Coqsrc}
 Compute test 4.
\end{Coqsrc}

\begin{Coqanswer}
  = {| round := 4; n2 := 2; n1 := 4; nh := 0 |}
     : state
\end{Coqanswer}

\begin{Coqsrc}
Compute test 5.
\end{Coqsrc}

\begin{Coqanswer}
  = {| round := 5; n2 := 2; n1 := 3; nh := 5 |}
     : state
\end{Coqanswer}

\begin{Coqsrc}
Compute test 2000.
\end{Coqsrc}

\begin{Coqanswer}
  = {| round := 2000; n2 := 1; n1 := 90; nh := 1102 |}
     : state
\end{Coqanswer}

The battle we are studying seems to be awfully long. Let us concentrate our
tests on some particular events : the states where $\texttt{nh}=0$.
From the value of \texttt{test 5},  it is obvious that at the 10-th round, the counter \texttt{nh} is equal to zero.

 \begin{Coqsrc}
 Compute test 10.
\end{Coqsrc}

\begin{Coqanswer}
    = {| round := 10; n2 := 2; n1 := 3; nh := 0 |}
     : state
 \end{Coqanswer}

@@

\begin{Coqsrc}
Compute test 22.
\end{Coqsrc}

\begin{Coqanswer}
 = {| round := 22; n2 := 2; n1 := 2; nh := 0 |}
     : state
\end{Coqanswer}

\begin{Coqsrc}
 Compute test 46.
\end{Coqsrc}

\begin{Coqanswer}
( = {| round := 46; n2 := 2; n1 := 1; nh := 0 |}
     : state
\end{Coqanswer}

\begin{Coqsrc}
Compute test 94.
\end{Coqsrc}

\begin{Coqanswer}
 = {| round := 94; n2 := 2; n1 := 0; nh := 0 |}
     : state
\end{Coqanswer}

@@

\begin{Coqsrc}
 Compute test 95.
\end{Coqsrc}

\begin{Coqanswer}
  = {| round := 95; n2 := 1; n1 := 95; nh := 0 |}
     : state
\end{Coqanswer}

@@

\begin{Coqsrc}
Definition doubleS (n : nat) := 2 * (S n).

Compute test (doubleS 95).
\end{Coqsrc}

\begin{Coqanswer}
 = {| round := 192; n2 := 1; n1 := 94; nh := 0 |}
     : state
\end{Coqanswer}

\begin{Coqsrc}
Compute test (iterate doubleS 2 95).
\end{Coqsrc}

\begin{Coqanswer}
  = {| round := 386; n2 := 1; n1 := 93; nh := 0 |}
     : state
\end{Coqanswer}

>>>>

\index{hydras}{Library Hydra!Predicates!round\_n}

\input{movies/snippets/Hydra_Definitions/roundNDef}

@@

\index{hydras}{Library Hydra!Predicates!round}
\label{sect:infix-round}

\input{movies/snippets/Hydra_Definitions/roundDef}

@@

\input{movies/snippets/Hydra_Definitions/roundPlus}

@@

\begin{exercise}
  Prove that if \texttt{$h$ -+-> $h'$}, then
  the height of $h'$ is less or equal than the height of $h$.

@@

\input{movies/snippets/Hydra_Definitions/BattleDef}

@@

\input{movies/snippets/Hydra_Definitions/freeDef}

@@

\input{movies/snippets/Hydra_Definitions/standardDef}

@@

\input{movies/snippets/Hydra_Definitions/battleRelDef}

The following property allows us to build battles by composition of smaller ones.

%% TODO display sub-goals when fixed

\input{movies/snippets/Hydra_Lemmas/battleTrans}

@@

\input{movies/snippets/BigBattle/hinitDef}

@@

\input{movies/snippets/BigBattle/L02}

@@

A first study with pencil and paper suggested us that, after three rounds, the hydra always looks like in figure~\vref{fig:hinit-plusn} (with a variable number of 
subtrees of height 1 or 0).
Thus, we introduce a few handy abbreviations.

\input{movies/snippets/BigBattle/Notations}

@@

\input{movies/snippets/BigBattle/L23L03}

@@

\input{movies/snippets/BigBattle/stateDef}

@@

\input{movies/snippets/BigBattle/nextDef}

@@

\input{movies/snippets/Iterates/iterateDef}

@@

\input{movies/snippets/BigBattle/testDefTests}

The battle we are studying looks to be awfully long. Let us concentrate our
tests on some particular events : the states where $\texttt{nh}=0$.
From the value of \texttt{test 5},  it is obvious that at the 10-th round, the counter \texttt{nh} is equal to zero.

\input{movies/snippets/BigBattle/smartTest}

@@

\input{movies/snippets/BigBattle/smartTestb}

@@

\input{movies/snippets/BigBattle/smartTestc}

@@

\input{movies/snippets/BigBattle/doubleS}
